### 2022-09-26 main/chapter_context.tex
add a bunch of citations for standard model

Mathematically, the Standard Model is described by a quantized gauge field
theory with the symmetry group
$SU\!(3)_C \times SU\!(2)_L \times U\!(1)_Y$,
meaning that its Lagrangian is invariant under local operations of group
elements on its fields in certain representations.

@@

The fermions of the Standard Model are listed in Table~\TODO{}.
Older Standard Models included only two fermion generations and no
neutrino masses~\cite{
wells2020discovery
},\TODO{cite more}
but a modern model must include both three generations and non-zero neutrino
masses to be compatible with experiment:
three fermion generations are directly established
\TODO{kaon mixing CPV, direct observation of tau/top/bottom, Z width},
as are neutrino masses from their flavour oscillations demonstrating that
they experience proper time~\TODO{cite}.

\TODO{fermion structure}
\TODO{three generations}
\TODO{rotation between weak and mass eigenstates: CKM matrix}

@@

\paragraph{Gravitation.}

\paragraph{Dark matter.}
\TODO{
primordial black holes,
WIMPs,
axions,
dust?
}
\TODO{combinations!}

\paragraph{Cosmology.}
\TODO{dark matter, cosmological constant, inflation, baryogenesis (matter-antimatter asymmetry)}
\TODO{note \lcdm}

@@

\paragraph{Aesthetics.}
\TODO{naturalness and fine-tuning}
\TODO{strong CP problem}
\TODO{hierarchy problem}

@@

>>>>

Mathematically, the Standard Model is described by a quantized gauge field
theory with the symmetry group
$SU\!(3)_C \!\times\! SU\!(2)_L \!\times\! U\!(1)_Y$,
meaning that its Lagrangian is invariant under local operations of group
elements on its fields in certain representations.

@@

The fermions of the Standard Model are listed in Table~\TODO{}.
% Glashow-Salam-Weinberg
Older Standard Models included only two fermion generations and no
neutrino masses~\cite{
wells2020discovery
},\TODO{cite more}
but a modern model must include both three generations and non-zero neutrino
masses to be compatible with experiment:
three fermion generations are directly established
\TODO{kaon mixing CPV~\cite{
cronin1964evidence,
kabayasji1973cpv
}, direct observation of tau/top/bottom~\cite{
perl1977evidence,
herb1977observation,
abachi1995observation
}, Z width~\cite{
lep2006precision
}
},
as are neutrino masses from their flavour oscillations demonstrating that
they experience proper time~\cite{
kamiokande1998measurement,
superk1998evidence,
superk1999measurement,
lsnd1998evidence
}.

\TODO{fermion structure}
\TODO{three generations}
\TODO{rotation between weak and mass eigenstates: CKM matrix~\cite{
cabibbo1963unitary,
kabayasji1973cpv
}}

@@

\TODO{mass models: dirac, Majorana, seesaw~\cite{
wells2020discovery,
king2003neutrino,
thomson2013modern
}}

\paragraph{Gravitation and cosmology.}

\TODO{dark matter: rotation curves~\cite{
begeman1991rotation,
garrett2010dark
}}

\TODO{
primordial black holes~\cite{bernard2019primordial},
WIMPs~\cite{jungman1996wimp},
axions~\cite{kim2008axions},
dust?~\cite{garrett2010dark}
}
\TODO{combinations!}

\TODO{
dark matter,
cosmological constant~\cite{riess1998observational},
inflation~\cite{bezrukov2008higgs},
baryogenesis (matter-antimatter asymmetry)
}
\TODO{note \lcdm}

@@

\paragraph{Aesthetics.}

\TODO{naturalness and fine-tuning~\cite{wells2020discovery}}
\TODO{strong CP problem}
\TODO{hierarchy problem}

@@

\TODO{higgs mass $\lesssim 130\,\eV[G]$~\cite{
quiros1997constraints,
wells2020discovery
}}

\TODO{primer~\cite{
martin2016primer
}}
